---
title: "Econ 1042 PS1"
author: "Owen Asnis"
date: "2023-02-01"
output: pdf_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(estimatr)
library(sandwich)
library(lmtest)
options(scipen = 5)

```

# Data Wrangling 

```{r data wrangling, message=FALSE}

kickers <- read_csv("kickers_v2.csv",
                    show_col_types = FALSE) %>%
  rename(., "ID" = "...1") %>%
  mutate(Grass = if_else(Grass == TRUE, 1, 0))

```

# Question 1

### a.)
```{r Q1a}

min <- min(kickers$Distance)
max <- max(kickers$Distance)
mean <- round(mean(kickers$Distance), 
              digits = 3)
median <- median(kickers$Distance)

sprintf("Minimum distance: %s yards.", min)
sprintf("Maximum distance: %s yards.", max)
sprintf("Mean distance: %s yards", mean)
sprintf("Median distance: %s yards", median)

```

### b.) The minimum distance (18 yards) isn't lower for two reasons: First, the kicker must kick the ball through the end zone (accounting for 10 yards) in addition to the distance from the ball to the goal line, and second the ball is snapped a distance of 7 yards to the holder, who holds the ball for the kicker. Therefore, an 18 yard field goal actually begins at the 1 yard line, but the total distance of the kick is 10 yards (the end zone) + 7 yards (the snap) + 1 yard (the distance to the goal line) = 18 yards. 

### c.)
```{r Q1c}

max_distance_kick <- kickers %>% 
  filter(Distance == max(Distance))

print(max_distance_kick)

```
### The special circumstance that explains the maximum is that it occured in the 30th minute of the game, the last minute of the 2nd quarter before halftime. Therefore, in this instance, the Oakland Raiders were forced to try to score, because the half was ending and possession is assigned based on the result of the coin toss after the 2nd quarter, and decided they had a better chance of scoring on a very long field goal rather than a hail mary. 

# Question 2

```{r Q2}

forty_to_fortyfive <- kickers %>%
  filter(Distance %in% (40:45))

forty_to_fortyfive_success <- round((mean(forty_to_fortyfive$Success) * 100), 
                                    digits = 3)

over_fortyfive <- kickers %>% 
  filter(Distance > 45)

over_fortyfive_success <- round((mean(over_fortyfive$Success) * 100), 
                            digits = 3)

sprintf("Kicks from 40 to 45 yards made: %s percent", forty_to_fortyfive_success)
sprintf("Kicks over 45 yards made: %s percent", over_fortyfive_success)

```

# Question 3

```{r Q3p1}

grass_only <- kickers %>% 
  filter(Grass == 1)

make_rate_grass <- round((mean(grass_only$Success) * 100), 
                         digits = 3)

turf_only <- kickers %>% 
  filter(Grass == 0)

make_rate_turf <- round((mean(turf_only$Success) * 100), 
                        digits = 3)

sprintf("Make rate on grass: %s percent", make_rate_grass)
sprintf("Make rate on turf: %s percent", make_rate_turf)

```
### The make rate was slightly higher on turf. 

```{r Q3p2}

m1 <- lm(Success ~ Grass, data = kickers)
coeftest(m1, vcov = vcovHC(m1))

```
### The difference is statistically significant at the 0.001 level. It could be the true effect of surface, because turf is consistantly a smoother and more solid surface, whereas grass is more imperfect with divits and inconsistencies. Therefore, in some ways, it seems like turf provides better conditions for kickers. However, there are clear counterarguments, including that grass might be better in rainy and snowy conditions, because it would be less slippery than turf.

# Question 4 

### a.)
```{r Q4a}

distance_surface_correlation <- round(cor(kickers$Distance, kickers$Grass), 
                                       digits = 3)

sprintf("Correlation between distance and surface: %s",
        distance_surface_correlation)

```
### Coaches might be more reluctant to attempt longer kicks on grass because of its imperfect texture and also potentially because many grass fields are outdoors where wind conditions can drastically influence kicking whereas many turf fields are indoors where these conditions aren't a factor.  

### b.)
```{r Q4b}

distance_makepc_correlation <- round(cor(kickers$Distance, kickers$Success), 
                                     digits = 3)

sprintf("Correlation between distance and make percentage: %s",
        distance_makepc_correlation)

```

# Question 5 

### a.) The formula for omitted variable bias (in English) is: Short-Form Regression Coefficient = Long-Form Regression Coefficient + Omitted Variable Regression Coefficient x Correlation Between Orginal Variables. 

### b.) Given (a), when you add distance when estimating the effect of a kick being on grass, you would expect that the coefficient for surface would become more negative, because distance is negatively correlated with make percentage and is the omitted variable. 
```{r Q5b}

m2 <- lm_robust(Success ~ Grass + Distance, data = kickers)
summary(m2)

```
### When the regression is run with distance, the estimate for Grass becomes more negative, as expected. Specifically, it was -0.01933 (see Question 3) and now is -0.01997. 

# Question 6

### a.) 
```{r Q6a}

m3 <- lm_robust(Success ~ Distance + Grass + ScoreDiff + GameMinute, 
                 data = kickers)
summary(m3)

```
### For every yard of distance added, the kick is roughly 1.2% less likely to be made. When the surface is grass, the kick is roughly 2% less likely to be made than if the surface was turf. For each point of score differential added, the kick is roughly 0.01% less likely to be made. And for each game minute gone by, the kick is 0.0004% more likely to be made. Therefore, it doesn't seem that kickers do better or worse late in the game and the score of the game doesn't seem to effect them, because both coefficients are so small that they are more likely to be caused by random noise.  

### b.)
```{r Q6b}

m4 <- lm_robust(Success ~ Distance + Grass + ScoreDiff + GameMinute + Kicker, 
                 data = kickers)
summary(m4)

```
### Kicker fixed effects corrects for the skill of each individual kicker. Adjusted R-squared increases from 0.1139 to 0.1205, which makes sense because this is a more complex model and seems to be better fitted. 

# Question 7 

### a.)
```{r Q7a}

new_data <- kickers %>%
  filter(Kicker == "Tucker", 
         ScoreDiff == -11, 
         Year == 2015, 
         GameMinute == 30)

tucker_lm_predict <- round((predict(m4, newdata = new_data) * 100), 
                           digits = 3)

sprintf("Probability of Justin Tucker making specified kick: %s percent", 
        tucker_lm_predict)

```

### b.)
```{r Q7b}

new_data

```
### The distance of the kick was only 30 yards and Justin Tucker is one of the best kickers in NFL history, so that prediction does seem reasonable to me. 

```{r Q7c}

average_lm_predict <- round((tucker_lm_predict - 0.10472919),
                            digits = 3)

sprintf("Probability of an average kicker making specified kick: %s percent", 
        average_lm_predict)

```

# Question 8 

### a.)
```{r Q8}

m5 <- glm(Success ~ Distance + Grass + ScoreDiff + GameMinute + Kicker,
          family = binomial(),
          data = kickers)

tucker_glm_predict <- round((predict(m5, newdata = new_data, 
                                     type = "response") * 100), 
                           digits = 3)

sprintf("Probability of Justin Tucker making specified kick: %s percent", 
        tucker_glm_predict)

```

### b.) The coefficients look so different for the logistic regression versus the OLS regression, because coefficients in logistic regressions show a value of log odds whereas coefficients in OLS regressions show how each interval of one variable changes the other variable (Source: "Interpreting Coefficients in Linear and Logistic Regression" by Jonathan Benton). 

# Question 9

### a.) 
```{r Q9a}

kickers %>% 
  group_by(Kicker) %>% 
  summarise(make_rate = mean(Success), 
            total_makes = sum(Success)) %>% 
  arrange(desc(make_rate)) %>% 
  filter(total_makes > 200) %>% 
  head(1)

```
### According to this analysis, Stephen Gostkowski was the best kicker in the NFL over this period. Gostkowski had the highest make percentage among kickers with at least 200 total makes in the dataset. In other words, he was consistently a high percentage kicker, and as a New England Patriots fan, this feels very right! 

```{r Q9a2}

kickers %>% 
  filter(Distance > 45) %>% 
  group_by(Kicker) %>% 
  summarise(total_makes = sum(Success), 
            over_fortyfive_percentage = mean(Success)) %>% 
  filter(total_makes > 50) %>% 
  arrange(desc(over_fortyfive_percentage)) %>% 
  head(1)
  
```
### According to this analysis, Adam Vinatieri was the best kicker in the NFL over this period. Vinatieri had the highest make percentage of attempts over 45 yeards among kickers with at least 50 total makes above 45 yards in the dataset. In other words, Vinatieri was consistantly a high percentage kicker from long range, and once again, as a Patriots fan, this feels right! 

### b.)
```{r Q9b}

kickers %>% 
  filter(Kicker == "Gostkowski") %>% 
  group_by(Year) %>% 
  summarise(make_rate = mean(Success))

```
### Overall, Gostkowki's make rate is fairly stable over time. There are some ebbs and flows, but generally he's hovering around the low 80% range to low 90% range. 

```{r Q9b2}

kickers %>% 
  filter(Distance > 45, 
         Kicker == "Vinatieri") %>% 
  group_by(Year) %>% 
  summarise(over_fortyfive_percentage = mean(Success))
  

```
### Vinatieri's make rate of attempts over 45 yards isn't stable over time. Specifically, between 2005-2009, Vinatieri's make rate for attempts over 45 yards falls below his rate for the whole dataset in 4 of 5 years, and between 2010-2014, his make rate is above his rate for the whole dataset in 4 of 5 years. Altogether, this shows that Vinatieri's make rate of attempts over 45 yards rose with more experience. 

# Question 10

```{r Q10}

simple <- kickers %>% 
  mutate(kick = 1) %>% 
  group_by(Kicker) %>% 
  summarise(make_rate = mean(Success), 
            total_kicks = sum(kick)) %>% 
  arrange(desc(make_rate)) %>% 
  mutate(total_kicks_squared = (total_kicks)^2)

m6 <- lm_robust(make_rate ~ total_kicks, 
                data = simple)
summary(m6)

m7 <- lm_robust(make_rate ~ total_kicks + total_kicks_squared, 
                data = simple)
summary(m7)

```
### The regression with a linear specification shows some support to the conjecture that kickers get better with experience in this dataset, as for ever additional kick a kicker performs, their make rate slightly increases. The regression with a quadratic specification shows similar support at first, but eventually, shows that kickers performance would decline. A way to interpret this is that kickers get better with experience at first, but their performance eventually declines with age. With this being said, these estimates certainly aren't perfect, as they don't consider many potential confounding variables and don't study how kickers' make rates change throughout this specific dataset. 

# Question 11 

### 1. Precipitation? - Precipitation would negatively bias the kicker, because it's harder to kick in rainy or snowy conditions. 
### 2. Centered Spot? - A centered spot would positively bias the kicker, because it's more of a straight shot for the kicker to make the attempt.  
### 3. Blocked? - A blocked kick would negatively bias the kicker, because even though the attempt wasn't made, it wasn't the kicker's fault (it was the lineman's fault). 

### Note: I worked with Ty Thabit on this problem set. 